A Survey on Ordered Binary Decision Diagrams

### 2015-09-14 paper.tex
Add full proof reference to BRYANT91
Finish integer multiplications part of Limitations

Bryant~\cite{BRYANT91} first generalized a function over $2n$ variables. The size
of an OBDD then can range from $2n$ up to $2^{n+1}-2$ for different variable
orderings. Now, analyzing $n$-bit integer multiplication, Bryant~\cite{BRYANT86}
already proved that for any variable ordering, there exists an $0 \le i \le 2n$
such that the function representing output $i$ requires an OBDD of size
$\Omega(1.09^n)$. Bryant himself~\cite{BRYANT91} showed that there is not much
of a difference between computing the single output $(n-1)$ or the entire product
of an $n$-bit integer multiplier. Bryant~\cite{BRYANT91} found the $n$-bit integer
multiplication to be especially hard, requiring an OBDD having at least $2^{n/8}$
nodes to compute the middle bit of the built product for any variable ordering.
Hence, the complexity of an OBDD representing integer multiplication is
exponential in $n$, where $n$ denotes the word size of the multiplied integers.

Nevertheless, researches put some effort in improving this lower bound. In 2001,
Woelfel~\cite{WOELFEL01} published a new bound lower bound as well as a first 
nontrivial upper bound for the size of OBDDs representing $n$-bit integer
multiplication by utilizing the concept of \textit{universal hashing}. \tbr

>>>>

Bryant~\cite{BRYANT91} first generalized a function over $2n$ variables. The size
of an OBDD then can range from $2n$ up to $2^{n+1}-2$ for different variable
orderings. Now, analyzing $n$-bit integer multiplication, Bryant~\cite{BRYANT86}
already proved that for any variable ordering, there exists an $0 \le i \le 2n$
such that the function representing output $i$ requires an OBDD of size
$\Omega(1.09^n)$. Bryant himself~\cite{BRYANT91} showed that there is not much
of a difference between computing the single output $(n-1)$ or the entire product
of an $n$-bit integer multiplier. Bryant~\cite{BRYANT91} found the $n$-bit integer
multiplication to be especially hard, requiring an OBDD having at least $2^{n/8}$
nodes to compute the middle bit of the built product for any variable ordering.
Hence, the complexity of an OBDD representing integer multiplication is
exponential in $n$, where $n$ denotes the word size of the multiplied integers.
One can read the full proof in~\cite{BRYANT91}.

Nevertheless, researches put some effort in improving this lower bound. In 2001,
Woelfel~\cite{WOELFEL01} published a new bound lower bound as well as a first 
nontrivial upper bound for the size of OBDDs representing $n$-bit integer
multiplication by utilizing the concept of \textit{universal hashing}. Therefore,
he utilized a universal family of hash functions to improve the exponential lower
bound on the OBDD-size of integer multiplication to
$2^{\lfloor\frac{n}{2}\rfloor}/96$. Moreover, Woelfel~\cite{WOELFEL01} proved
the upper bound for the OBDD-size of computing the middle bit of an $n$-bit
multiplication to be $7/3 \cdot 2^{\frac{4 \cdot n}{3}}$.

Nevertheless, all lower bounds for $n$-bit integer multiplication are exponential
in the number of nodes. Especially because of the high practical relevance of
$n$-bit integer multiplication this is one of the main drawbacks of the OBDD
representation~\cite{BRYANT86, BRYANT91, WOELFEL01}. \tbr
